---
title: "county_risk_score"
author: "Chen Chen"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
# Set up
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)); getwd();
source("../base/global.R")
x.date <- as.character(Sys.Date(), format = '%Y%m%d')

knitr::opts_chunk$set(echo = TRUE)

normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
```

## STEP 1
## READ ACS data

```{r ACS data}

x.state.pop = dbGetQuery(coririsi_layer, "SELECT geoid, st_total_population_2018 FROM attr_state_full")

x.layer.name <- "attr_county_health"
x.layer.table = st_read(coririsi_layer, x.layer.name) %>% 
  left_join(x.state.pop, by = c('geoid_st' = 'geoid')) %>% 
  dplyr::mutate(svi_socioeconomic = ifelse(svi_socioeconomic == -999, mean(svi_socioeconomic, na.rm = T), svi_socioeconomic), 
                total_estimated_bed_40_mins_100k = total_estimated_bed_40_mins*100000/total_population_2018,
                total_staff_100k = total_staff*100000/total_population_2018,
                icuover_max_needed_100k = icuover_max_needed*100000/st_total_population_2018 * (total_population_2018/st_total_population_2018)) %>% 
  drop_na(pct_65_over_2018, svi_socioeconomic)

names(x.layer.table)
```

```{r ACS data}
x.layer.table.rg = x.layer.table %>% 
  st_drop_geometry() %>%
  dplyr::select(geoid, name, st_stusps, total_estimated_bed_40_mins, total_staff, pct_65_over_2018, svi_socioeconomic, icuover_max_needed, total_population_2018) %>% 
  group_by(st_stusps) %>% 
  dplyr::mutate(st_total_population_2018 = sum(total_population_2018)) %>%
  ungroup() %>%
  data.frame() %>%
  dplyr::mutate(total_estimated_bed_40_mins_100k = total_estimated_bed_40_mins*100000/total_population_2018,
                total_staff_100k = total_staff*100000/total_population_2018,
                icuover_max_needed_100k = icuover_max_needed*100000/st_total_population_2018 * (total_population_2018/st_total_population_2018))


x.layer.table.rg$norm_total_estimated_bed_40_mins_100k = normalize(x.layer.table.rg[,"total_estimated_bed_40_mins_100k"])

x.layer.table.rg$norm_total_staff_100k = normalize(x.layer.table.rg[,"total_staff_100k"])

x.layer.table.rg$norm_pct_65_over_2018 = normalize(x.layer.table.rg[,"pct_65_over_2018"])

x.layer.table.rg$norm_svi_socioeconomic = normalize(x.layer.table.rg[,"svi_socioeconomic"])

x.layer.table.rg$norm_icuover_max_needed_100k = normalize(x.layer.table.rg[,"icuover_max_needed_100k"])

x.layer.table.rg = x.layer.table.rg %>% 
  dplyr::mutate(pctl_total_estimated_bed_40_mins_100k = ntile(norm_total_estimated_bed_40_mins_100k, 100),
                pctl_pct_65_over_2018 = ntile(-norm_pct_65_over_2018, 100),
                pctl_total_staff_100k = ntile(norm_total_staff_100k, 100),
                pctl_icuover_max_needed_100k = ntile(-norm_icuover_max_needed_100k, 100),
                pctl_svi_socioeconomic = ntile(-norm_svi_socioeconomic, 100)) %>% 
    dplyr::select(geoid, name, st_stusps, everything(), starts_with("norm_"), starts_with("pctl_"))

write.csv(x.layer.table.rg, "./output/county_health_factor_sets.csv", row.names = F)

names(x.layer.table.rg)
```


## STEP 2
## DATA EXPLORATION


```{r plots}
#Beds
hist(x.layer.table[,"total_estimated_bed_40_mins"], breaks=50)
hist(x.layer.table[,"total_icu_bed"], breaks=50)

#Staff
hist(x.layer.table[,"total_staff"], breaks=50)

#Socio-economic
hist(x.layer.table[,"svi_socioeconomic"], breaks=50)
# summary(x.layer.table[,"svi_socioeconomic"])
#Demographics
hist(x.layer.table[,"pct_65_over_2018"], breaks=50)
summary(x.layer.table[,"pct_65_over_2018"])
```





## Beds Pillar 

```{r housing index data, echo=FALSE}
beds_pillar = x.layer.table[, c("geoid", "name", "st_stusps" ,"total_estimated_bed_40_mins", "total_estimated_bed_40_mins_100k")]

# Make plots
plot_qq(beds_pillar)

beds_pillar %>%
  as.data.frame() %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free", nrow = 5) +
    geom_histogram()



```


```{r beds pillar method, echo=FALSE}

# Assign weights
weight1 = 1
weight2 = 1
weight3 = 1
weight4 = 1
weight5 = 1
weight6 = 1
weight7 = 1

# Calculate the composite index
beds_pillar = beds_pillar %>%
  dplyr::mutate(# Composite Index 1 uses percentile
                bed_score_1 = ntile(total_estimated_bed_40_mins_100k, 100),
                # Composite Index 2 uses rescale
                bed_score_2 = as.integer(rescale(total_estimated_bed_40_mins_100k, to = c(0, 100))))

beds_pillar_plot = beds_pillar %>%
  dplyr::filter(st_stusps %in% states.cont)

# Skim the distribution of the composite index
hist(beds_pillar$bed_score_1)
hist(beds_pillar$bed_score_2)


# Quick Map using two indexes
tm_shape(beds_pillar_plot) + 
  tm_fill(col = "bed_score_1", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)

tm_shape(beds_pillar_plot) + 
  tm_fill(col = "bed_score_2", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)



```



## Staff Pillar 

```{r housing index data, echo=FALSE}
staff_pillar = x.layer.table[, c("geoid", "name", "st_stusps" ,"total_staff", "total_staff_100k")]

# Make plots
plot_qq(staff_pillar)

staff_pillar %>%
  as.data.frame() %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free", nrow = 5) +
    geom_histogram()



```


```{r housing index method, echo=FALSE}

# Assign weights
weight1 = 1
weight2 = 1
weight3 = 1
weight4 = 1
weight5 = 1
weight6 = 1
weight7 = 1

# Calculate the composite index
staff_pillar = staff_pillar %>%
  dplyr::mutate(# Composite Index 1 uses percentile
                staff_score_1 = ntile(total_staff_100k, 100),
                # Composite Index 2 uses rescale
                staff_score_2 = as.integer(rescale(total_staff_100k, to = c(0, 100))))

staff_pillar_plot = staff_pillar %>%
  dplyr::filter(st_stusps %in% states.cont)

# Skim the distribution of the composite index
hist(staff_pillar$staff_score_1)
hist(staff_pillar$staff_score_2)


# Quick Map using two indexes
tm_shape(staff_pillar_plot) + 
  tm_fill(col = "staff_score_1", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)

tm_shape(staff_pillar_plot) + 
  tm_fill(col = "staff_score_2", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)



```


## Demographic Pillar 

```{r housing index data, echo=FALSE}
dem_pillar = x.layer.table[, c("geoid", "name", "st_stusps" ,"pct_65_over_2018")]

# Make plots
plot_qq(dem_pillar)

dem_pillar %>%
  as.data.frame() %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free", nrow = 5) +
    geom_histogram()



```


```{r housing index method, echo=FALSE}

# Assign weights
weight1 = 1
weight2 = 1
weight3 = 1
weight4 = 1
weight5 = 1
weight6 = 1
weight7 = 1

# Calculate the composite index
dem_pillar = dem_pillar %>%
  dplyr::mutate(# Composite Index 1 uses percentile
                dem_score_1 = ntile(-pct_65_over_2018, 100),
                # Composite Index 2 uses rescale
                dem_score_2 = as.integer(rescale(-pct_65_over_2018, to = c(0, 100))))

dem_pillar_plot = dem_pillar %>%
  dplyr::filter(st_stusps %in% states.cont)

# Skim the distribution of the composite index
hist(dem_pillar$dem_score_1)
hist(dem_pillar$dem_score_2)


# Quick Map using two indexes
tm_shape(dem_pillar_plot) + 
  tm_fill(col = "dem_score_1", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)

tm_shape(dem_pillar_plot) + 
  tm_fill(col = "dem_score_2", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)



```



## IHME Pillar 

```{r housing index data, echo=FALSE}
proj_pillar = x.layer.table[, c("geoid", "name", "st_stusps" ,"icuover_max_needed", "icuover_max_needed_100k")]

# Make plots
plot_qq(proj_pillar)

proj_pillar %>%
  as.data.frame() %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free", nrow = 5) +
    geom_histogram()
```

```{r housing index method, echo=FALSE}

# Assign weights
weight1 = 1
weight2 = 1
weight3 = 1
weight4 = 1
weight5 = 1
weight6 = 1
weight7 = 1

# Calculate the composite index
proj_pillar = proj_pillar %>%
  dplyr::mutate(# Composite Index 1 uses percentile
                proj_score_1 = ntile(-icuover_max_needed_100k, 100),
                # Composite Index 2 uses rescale
                proj_score_2 = as.integer(rescale(-icuover_max_needed_100k, to = c(0, 100))))

proj_pillar_plot = proj_pillar %>%
  dplyr::filter(st_stusps %in% states.cont)

# Skim the distribution of the composite index
hist(proj_pillar$proj_score_1)
hist(proj_pillar$proj_score_2)


# Quick Map using two indexes
tm_shape(proj_pillar_plot) + 
  tm_fill(col = "proj_score_1", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)

tm_shape(proj_pillar_plot) + 
  tm_fill(col = "proj_score_2", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)



```


## Socio-eco Pillar 
```{r housing index data, echo=FALSE}
se_pillar = x.layer.table[, c("geoid", "name", "st_stusps" ,"svi_socioeconomic")]

# Make plots
plot_qq(se_pillar)

se_pillar %>%
  as.data.frame() %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free", nrow = 5) +
    geom_histogram()

```


```{r housing index method, echo=FALSE}

# Assign weights
weight1 = 1
weight2 = 1
weight3 = 1
weight4 = 1
weight5 = 1
weight6 = 1
weight7 = 1

# Calculate the composite index
se_pillar = se_pillar %>%
  dplyr::mutate(# Composite Index 1 uses percentile
                se_score_1 = ntile(-svi_socioeconomic, 100))

se_pillar_plot = se_pillar %>%
  dplyr::filter(st_stusps %in% states.cont)

# Skim the distribution of the composite index
hist(se_pillar$se_score_1)


# Quick Map using two indexes
tm_shape(se_pillar_plot) + 
  tm_fill(col = "se_score_1", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)




```


## Overall composite index

``````{r overall, echo=FALSE}

all_index = x.layer.table[, c('geoid', 'name', 'st_stusps', 'icuover_max_date', 'icu_bed_max_date')] %>% 
  left_join(beds_pillar %>% data.frame %>% dplyr::select(-geom) , by= c('geoid', 'name', 'st_stusps')) %>%
  left_join(staff_pillar %>% data.frame %>% dplyr::select(-geom) , by= c('geoid', 'name', 'st_stusps')) %>%
  left_join(dem_pillar %>% data.frame %>% dplyr::select(-geom), by= c('geoid', 'name', 'st_stusps')) %>%
  left_join(se_pillar %>% data.frame %>% dplyr::select(-geom), by= c('geoid', 'name', 'st_stusps')) %>% 
  left_join(proj_pillar %>% data.frame %>% dplyr::select(-geom), by= c('geoid', 'name', 'st_stusps')) %>% 
    dplyr::mutate(prep_score = ( bed_score_1 + staff_score_1 + dem_score_1 + se_score_1 + proj_score_1) / 5,
                  prep_score_1 = ntile(prep_score, 100),
                  prep_score_2 = as.integer(rescale(prep_score, to = c(0, 100)))) %>% 
  dplyr::mutate(prep_level = case_when(prep_score_1 <= 20 ~ "Extremely Low Preparedness", 
                                       prep_score_1 > 20 & prep_score_1 <= 40 ~ "Low Preparedness", 
                                       prep_score_1 > 40 & prep_score_1 <= 60 ~ "Medium Preparedness", 
                                       prep_score_1 > 60 & prep_score_1 <= 80 ~ "High Preparedness", 
                                       prep_score_1 > 80 ~ "Well-prepared")) %>% 
  dplyr::mutate(svi_socioeconomic = round(svi_socioeconomic * 100)/100,
                pct_65_over_2018 = round(pct_65_over_2018 * 10)/10,
                total_estimated_bed_40_mins_100k = round(total_estimated_bed_40_mins_100k),
                total_staff_100k = round(total_staff_100k),
                icuover_max_needed_100k = round(icuover_max_needed_100k * 100)/100)


names(all_index)

hist(all_index$prep_score)
hist(all_index$prep_score_1)
hist(all_index$prep_score_2)

all_index_plot = all_index %>%
  dplyr::filter(st_stusps %in% states.cont)

# Quick Map using two indexes
tm_shape(all_index_plot) + 
  tm_fill(col = "prep_score", 
          palette = "PuBu", 
          breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
  tm_legend(outside = TRUE)


write_layer(all_index, layer.name = "county_preparedness_score", layer.type="pg", db=T, fs=T, , ngacarto=T, ngacarto.overwrite = T, new.server = T, new.server.overwrite = T)



```
